### 2021-10-21 reviews/01_ELE/01_draft_revised.tex
🚮 add the diff file

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen
2016). First, networks within a location are difficult to sample
properly (Jordano 2016a, b), resulting in a widespread ``Eltonian
shortfall'' (Hortal \emph{et al.} 2015), \emph{i.e.} a lack of knowledge
about inter and intra specific relationships. This first challenge has
been, in large part, addressed by the recent emergence of a suite of
methods aiming to predict interactions within \emph{existing} networks,
many of which are reviewed in Strydom \emph{et al.} (2021a). Second,
recent analyses based on collected data (Poisot \emph{et al.} 2021a) or
metadata (Cameron \emph{et al.} 2019) highlight that ecological networks
are currently studied in a biased subset of space and bioclimates, which
impedes our ability to generalize any local understanding of network
structure. Meaning that, although the framework to address
incompleteness \emph{within} networks exists, there would still be
regions for which, due to a \emph{lack} of local interaction data, we
are unable to infer potential species interactions. Having a general
solution for inferring a \emph{plausible} metaweb (despite the
unavailability of interaction data) could be the catalyst for
significant breakthroughs in our ability to start thinking about species
interaction networks over large spatial scales.

Here, we present a general method for the transfer learning of network
representations, relying on the similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang 2010; Torrey
\& Shavlik 2010). In this instance, we solve the problem of predicting
trophic interactions between species, based on knowledge extracted from
another species pool for which interactions are known by using
phylogenetic structure as a medium for transfer. This allows us to
construct a \emph{probabilistic} metaweb for a community for which we
have \emph{no} prior trophic interaction data for the desired species
pool. Our methodology is outlined in fig.~\ref{fig:concept}, where we
provide an illustration based on learning the embedding of a metaweb of
trophic interactions for European mammals (known interactions; Maiorano
\emph{et al.} 2020b, a) and, based on phylogenetic relationships between
mammals globally (\emph{i.e.}, phylogenetic tree Upham \emph{et al.}
2019), infer a metaweb for the Canadian mammalian species pool
(interactions are treated as unknown in this instance).

@@

There is a plurality of measures of species similarities that can be
used for metaweb reconstruction (see \emph{e.g.} Morales-Castilla
\emph{et al.} 2015); however, phylogenetic proximity has several
desirable properties when working at large scales. Gerhold \emph{et al.}
(2015) made the point that phylogenetic signal captures diversification
of characters (large macro-evolutionary process), but not necessarily
community assembly (fine ecological process); Dormann \emph{et al.}
(2010) previously found very similar conclusions. Interactions tend
reflect a phylogenetic signal because they have a conserved pattern of
evolutionary convergence that encompasses a wide range of ecological and
evolutionary mechanisms (Cavender-Bares \emph{et al.} 2009; Mouquet
\emph{et al.} 2012), and - most importantly - retain this signal even
when it is not detectable at the community scale (Hutchinson \emph{et
al.} 2017; Poisot \& Stouffer 2018). Finally, species interactions at
macro-ecological scales seem to respond mostly to macro-evolutionary
processes (Price 2003); which is evidenced by the presence of conserved
backbones in food webs (Dalla Riva \& Stouffer 2016), strong
evolutionary signature on prey choice (Stouffer \emph{et al.} 2012), and
strong phylogenetic signature in food web intervality (Eklöf \& Stouffer
2016). Phylogenetic reconstruction has also previously been used within
the context of ecological networks, namely understanding ancestral
plant-insect interactions (Braga \emph{et al.} 2021). Taken together,
these considerations suggest that phylogenies can reliably be used to
transfer knowledge on species interactions.

Our case study shows that phylogenetic transfer learning is indeed an
effective approach to predict the Canadian mammalian metaweb. This
showcases that although the components (species) that make up the
Canadian and European communities may be \emph{minimally} shared, if the
medium (proxy space) selected in the transfer step is biologically
plausible, we can still effectively learn from the known network and
make biologically relevant predictions of interactions. It should be
reiterated that the framework presented in fig.~\ref{fig:concept} is
amenable to changes; notably, the measure of similarity may not be
phylogeny, and can be replaced by information on foraging (Beckerman
\emph{et al.} 2006), cell-level mechanisms (Boeckaerts \emph{et al.}
2021), or a combination of traits and phylogenetic structure (Stock
2021).

@@

We use data from the European metaweb assembled by Maiorano \emph{et
al.} (2020b), following the definition of the metaweb first introduced
by Dunne (2006), \emph{i.e.} an inventory of all possible interactions
within species from a spatially delimited pool. Notably the metaweb is
not a prediction of the food web at any specific locale within the
frontiers of the species pool -- in fact, these local food webs are
expected to have a subset of both the species and the interactions of
their metaweb (Poisot \emph{et al.} 2012). This being said, as the
metaweb represents the total of functional, phylogenetic, and
macroecological processes (Morales-Castilla \emph{et al.} 2015), it is
thus still worthy of ecological attention. We deduced the subgraph
corresponding to all mammals by matching species names in the original
network to the GBIF taxonomic backbone (GBIF Secretariat 2021) and
retaining all those who matched to mammals. This serves a dual purpose
1) to extract only mammals from the European network and 2) to match and
standardize species names when aggregating the different data sources
further downstream (which is an important consideration when combining
datasets (Grenié \emph{et al.} 2021)). All nodes had valid matches to
GBIF at this step, and so this backbone is used for all name
reconciliation steps as outlined below.

@@

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the IUCN checklist, and selected the terrestrial and
semi-aquatic species (this corresponds to the same selection that was
applied by Maiorano \emph{et al.} (2020b) in the European metaweb). The
IUCN names were, as previously, reconciled against GBIF to have an exact
match to the taxonomy.

@@

@@

The entire pipeline is implemented in \emph{Julia} 1.6 (Bezanson
\emph{et al.} 2017) and is available under the permissive MIT License at
\href{https://osf.io/2zwqm/}{\texttt{https://osf.io/2zwqm/}}. The
taxonomic cleanup steps are done using \texttt{GBIF.jl} (Dansereau \&
Poisot 2021). The network embedding and analysis is done using
\texttt{EcologicalNetworks.jl} (Poisot \emph{et al.} 2019; Banville
\emph{et al.} 2021). The phylogenetic simulations are done using
\texttt{PhyloNetworks.jl} (Solís-Lemus \emph{et al.} 2017) and
\texttt{Phylo.jl} (Reeve \emph{et al.} 2016). A complete
\texttt{Project.toml} file specifying the full tree of dependencies is
available alongside the code. This material also includes a fully
annotated copy of the entire code required to run this project
(describing both the intent of the code and discussing some technical
implementation details), a vignette for every step of the process, and a
series of Jupyter notebooks with the text and code. The pipeline can be
executed on a laptop in a matter of minutes, and therefore does not
require extensive computational power.

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy \emph{et al.} 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer 2016). In addition, recent advances show that the
latent variables produced this way can be used to predict \emph{de novo}
network edges (\emph{i.e.} interactions; Runghen \emph{et al.} 2021).

@@

Because RDPG relies on matrix multiplication, the higher dimensions
essentially serve to make specific interactions converge towards 0 or 1;
therefore, for reasonably low ranks, there is no guarantee that the
values in the reconstructed network will be within the unit range. In
order to determine what constitutes an appropriate threshold for
probability, we performed the RDPG approach on the European metaweb, and
evaluated the probability threshold by treating this as a binary
classification problem, specifically assuming that both 0 and 1 in the
European metaweb are all true. Given the methodological details given in
Maiorano \emph{et al.} (2020b) and O'Connor \emph{et al.} (2020), this
seems like a reasonable assumption, although one that does not hold for
all metawebs. We used the thresholding approach presented in Poisot
\emph{et al.} (2021b), and picked a cutoff that maximized Youden's \(J\)
statistic (a measure of the informedness (trust) of predictions; Youden
(1950)); the resulting cutoff was 0.22, and gave an accuracy above 0.99.

@@

In order to transfer the knowledge from the European metaweb to the
Canadian species pool, we performed ancestral character estimation using
a Brownian motion model, which is a conservative approach in the absence
of strong hypotheses about the nature of phylogenetic signal in the
network decomposition (Litsios \& Salamin 2012). This uses the estimated
feature vectors for the European mammals to create a state
reconstruction for all species (conceptually something akin to a
trait-based mammalian phylogeny using generality and vulnerability
traits) and allows us to impute the missing (latent) trait data for the
Canadian species that are not already in the European network; as we are
focused on predicting contemporary interactions, we only retained the
values for the tips of the tree. We assumed that all traits (\emph{i.e.}
the feature vectors for the left and right subspaces) were independent,
which is a reasonable assumption as every trait/dimension added to the
t-SVD has an \emph{additive} effect to the one before it. Note that the
Upham \emph{et al.} (2019) tree itself has some uncertainty associated
to inner nodes of the phylogeny. In this case study, we have decided to
not propagate this uncertainty, as it would complexify the process. The
Brownian motion algorithm returns the \emph{average} value of the trait,
and its upper and lower bounds. Because we do not estimate other
parameters of the traits' distributions, we considered that every
species trait is represented as a uniform distribution between these
bounds; in a situation where the algorithm would return point values for
all simulations, one could in theory either estimate the parameters of a
distribution for each tip, or draw randomly from the outputs. In all
cases, the inferred left and right sub-spaces for the Canadian species
pool (\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\)) have entries that
are distributions, representing the range of values for a given species
at a given dimension.

@@

@@

As Herbert (1965) rightfully pointed out, ``{[}y{]}ou can't draw neat
lines around planet-wide problems''; in this regard, our approach must
contend with two interesting problems. The first is the limit of the
metaweb to embed and transfer. If the initial metaweb is too narrow in
scope, notably from a taxonomic point of view, the chances of finding
another area with enough related species to make a reliable inference
decrease. This is notably true if the metaweb is assembled in an area
with mostly endemic species. Conversely, the metaweb should be reliably
filled, which assumes that the \(S^2\) interactions in a pool of \(S\)
species have been examined, either through literature surveys or expert
elicitation. The second problem is to determine which area should be
used to infer the new metaweb in, as this determines the species pool
that must be used. In our application, we focused on the mammals of
Canada. The upside of this approach is that information at the country
level is likely to be required by policy makers and stakeholders for
their biodiversity assessment, as each country tends to set goals at the
national level (Buxton \emph{et al.} 2021) for which quantitative
instruments are designed (Turak \emph{et al.} 2017), with specific
strategies often enacted at smaller scales (Ray \emph{et al.} 2021). Yet
these national divisions, in large parts of the world, reflect nothing
except for the legacy of settler colonialism, and operating under them
must be done under the clear realization that they contributed to the
ongoing biodiversity crisis (Adam 2014), can reinforce environmental
injustice (Choudry 2013; Domínguez \& Luoma 2020), and on Turtle Island
especially, will probably end up being replaced by Indigenous principles
of land management (Eichhorn \emph{et al.} 2019; No'kmaq \emph{et al.}
2021).

@@

\leavevmode\hypertarget{ref-Arsov2019NetEmb}{}%
Arsov, N. \& Mirceva, G. (2019). Network Embedding: An Overview.
\emph{arXiv:1911.11726 {[}cs, stat{]}}.

@@

\leavevmode\hypertarget{ref-Brose2006ConRes}{}%
Brose, U., Jonsson, T., Berlow, E.L., Warren, P., Banasek-Richter, C.,
Bersier, L.-F., \emph{et al.} (2006). ConsumerResource Body-Size
Relationships in Natural Food Webs. \emph{Ecology}, 87, 2411--2417.

@@

\leavevmode\hypertarget{ref-Cai2017ComSur}{}%
Cai, H., Zheng, V.W. \& Chang, K.C.-C. (2017). A Comprehensive Survey of
Graph Embedding: Problems, Techniques and Applications. \emph{arXiv
preprint arXiv:1709.07604}.

@@

\leavevmode\hypertarget{ref-Cirtwill2019QuaFra}{}%
Cirtwill, A.R., Eklf, A., Roslin, T., Wootton, K. \& Gravel, D. (2019).
A quantitative framework for investigating the reliability of empirical
network construction. \emph{Methods in Ecology and Evolution}, 0.

@@

@@

\leavevmode\hypertarget{ref-Dansereau2021SimJl}{}%
Dansereau, G. \& Poisot, T. (2021). SimpleSDMLayers.jl and GBIF.jl: A
Framework for Species Distribution Modeling in Julia. \emph{Journal of
Open Source Software}, 6, 2872.

@@

@@

\leavevmode\hypertarget{ref-Herbert1965Dun}{}%
Herbert, F. (1965). \emph{Dune}. First. Chilton Book Company,
Philadelphia.

@@

\leavevmode\hypertarget{ref-Maiorano2020DatTet}{}%
Maiorano, L., Montemaggiori, A., Ficetola, G.F., O'Connor, L. \&
Thuiller, W. (2020a). Data from: Tetra-EU 1.0: A species-level trophic
meta-web of European tetrapods.

\leavevmode\hypertarget{ref-Maiorano2020TetEu}{}%
Maiorano, L., Montemaggiori, A., Ficetola, G.F., O'Connor, L. \&
Thuiller, W. (2020b). TETRA-EU 1.0: A species-level trophic metaweb of
European tetrapods. \emph{Global Ecology and Biogeography}, 29,
1452--1457.

@@

@@

\leavevmode\hypertarget{ref-OConnor2020UnvFoo}{}%
O'Connor, L.M.J., Pollock, L.J., Braga, J., Ficetola, G.F., Maiorano,
L., Martinez-Almoyna, C., \emph{et al.} (2020). Unveiling the food webs
of tetrapods across Europe through the prism of the Eltonian niche.
\emph{Journal of Biogeography}, 47, 181--192.

@@

@@

\leavevmode\hypertarget{ref-Poisot2012DisSpe}{}%
Poisot, T., Canard, E., Mouillot, D., Mouquet, N. \& Gravel, D. (2012).
The dissimilarity of species interaction networks. \emph{Ecology
Letters}, 15, 1353--1361.

@@

>>>>

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen
2016). First, networks within a location are difficult to sample
properly (Jordano 2016a, b), resulting in a widespread ``Eltonian
shortfall'' (Hortal \emph{et al.} 2015), \emph{i.e.} a lack of knowledge
about inter and intra specific relationships. This first challenge has
been, in large part, addressed by the recent emergence of a suite of
methods aiming to predict interactions within \emph{existing} networks,
many of which are reviewed in Strydom \emph{et al.} (2021a). Second,
recent analyses based on collected data (Poisot \emph{et al.} 2021a) or
metadata (Cameron \emph{et al.} 2019) highlight that ecological networks
are currently studied in a biased subset of space and bioclimates, which
impedes our ability to generalize any local understanding of network
structure. Meaning that, although the framework to address
incompleteness \emph{within} networks exists, there would still be
regions for which, due to a \emph{lack} of local interaction data, we
are unable to infer potential species interactions. Having a general
solution for inferring \emph{plausible} interactions (despite the
unavailability of interaction data) could be the catalyst for
significant breakthroughs in our ability to start thinking about species
interaction networks over large spatial scales. In a recent overview of
the field of ecological network prediction, Strydom \emph{et al.}
(2021a) identified two challenges of interest to the prediction of
interactions at large scales. First, there is a relative scarcity of
relevant data in most places globally -- paradoxically, this restricts
our ability to infer interactions to locations where inference is
perhaps the least required; second, accurate predictions often demand
accurate predictors, and the lack of methods that can leverage small
amount of data is a serious impediment to our predictive ability
globally.

Here, we present a general method to recommend possible trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang 2010; Torrey
\& Shavlik 2010). In this instance, we solve the problem of predicting
trophic interactions between species, based on knowledge extracted from
another species pool for which interactions are known by using
phylogenetic structure as a medium for transfer. There is a plurality of
measures of species similarities that can be used for metaweb
reconstruction (see \emph{e.g.} Morales-Castilla \emph{et al.} 2015);
however, phylogenetic proximity has several desirable properties when
working at large scales. Gerhold \emph{et al.} (2015) made the point
that phylogenetic signal captures diversification of characters (large
macro-evolutionary process), but not necessarily community assembly
(fine ecological process); Dormann \emph{et al.} (2010) previously found
very similar conclusions. Interactions tend reflect a phylogenetic
signal because they have a conserved pattern of evolutionary convergence
that encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares \emph{et al.} 2009; Mouquet \emph{et al.} 2012), and -
most importantly - retain this signal even when it is not detectable at
the community scale (Hutchinson \emph{et al.} 2017; Poisot \& Stouffer
2018). Finally, species interactions at macro-ecological scales seem to
respond mostly to macro-evolutionary processes (Price 2003); which is
evidenced by the presence of conserved backbones in food webs (Dalla
Riva \& Stouffer 2016), strong evolutionary signature on prey choice
(Stouffer \emph{et al.} 2012), and strong phylogenetic signature in food
web intervality (Eklöf \& Stouffer 2016). Phylogenetic reconstruction
has also previously been used within the context of ecological networks,
namely understanding ancestral plant-insect interactions (Braga \emph{et
al.} 2021). Taken together, these considerations suggest that
phylogenies can reliably be used to transfer knowledge on species
interactions.

@@

Our methodology is outlined in fig.~\ref{fig:concept}, where we provide
an illustration based on learning the embedding of a metaweb of trophic
interactions for European mammals (Maiorano \emph{et al.} 2020; known
interactions; \textbf{Maiorano2020TetEu?}) and, based on phylogenetic
relationships between mammals globally (\emph{i.e.}, phylogenetic tree
Upham \emph{et al.} 2019), infer a metaweb for the Canadian mammalian
species pool (interactions are treated as unknown in this instance).
Following the definition of Dunne (2006), a metaweb is a network
analogue to the concept of a regional species pool. Specifically, a
metaweb is an inventory of all \emph{possible} interactions within
species likely to occurr within a spatially delimited area (the network
\(\gamma\)-diversity, in a sense). The metaweb is, therefore, \emph{not}
a prediction of the food web at any specific locale within the frontiers
of the spatial area it recovers, and will in fact have a different
structure (notably by having a larger connectance; see \emph{e.g.}
\textbf{Wood2015EffSpa?}). These local food webs are expected to be a
subset of both the species and the interactions of their metaweb, and
have been called ``metaweb realizations'' (Poisot \emph{et al.} 2015).
The difference between a food web at a specific location and the metaweb
has to do with a variety of mechanisms, including species co-occurrence,
local environmental conditions, and local distribution of functional
traits. Nevertheless, the metaweb represents the total of functional,
phylogenetic, and macroecological processes (Morales-Castilla \emph{et
al.} 2015), and therefore still holds valuable ecological information.
Because the metaweb can be down-sampled given appropriate knowledge of
local species composition (the network \(\alpha\)-diversity, in a
sense), it is possible to infer what may drive the structure of food
webs at finer spatial scales. This has been done for example for
tree-gallers-parasitoid systems (\textbf{Gravel2018BriElt?}), fish
trophic interactions (Albouy \emph{et al.} 2019), tetrapods trophic
interactions (O'Connor \emph{et al.} 2020), and crop-pests networks
(Grünig \emph{et al.} 2020). Whereas the original metaweb definition was
based on presence/absence, we focus on \emph{probabilistic} metawebs;
not only does our method recommed interactions that may exist, it gives
each interaction a score that is mathematically equivalent to the chance
of success of a Bernoulli trial (see \emph{e.g.} Poisot \emph{et al.}
2016), which allows properly weigh interactions as a function of how
likely they are.

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the result, when validated against
known but fracational data of trophic interactions between Canadian
mammals, our model achieves a predictive accuracy of approx. 91\%. It
should be reiterated that the framework presented in
fig.~\ref{fig:concept} is amenable to changes; notably, the measure of
similarity may not be phylogeny, and can be replaced by information on
foraging (Beckerman \emph{et al.} 2006), cell-level mechanisms
(Boeckaerts \emph{et al.} 2021), or a combination of traits and
phylogenetic structure (Stock 2021).

@@

We use data from the European metaweb assembled by
(\textbf{Maiorano2020TetEu?}).

\textbf{TANYA TODO} description of the data

We induced the subgraph corresponding to all mammals by matching species
names in the original network to the GBIF taxonomic backbone (GBIF
Secretariat 2021) and retaining all those who matched to mammals. This
serves a dual purpose 1) to extract only mammals from the European
network and 2) to match and standardize species names when aggregating
the different data sources further downstream (which is an important
consideration when combining datasets (Grenié \emph{et al.} 2021)). All
nodes had valid matches to GBIF at this step, and so this backbone is
used for all name reconciliation steps as outlined below.

@@

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the IUCN checklist, and selected the terrestrial and
semi-aquatic species (this corresponds to the same selection that was
applied by (\textbf{Maiorano2020TetEu?}) in the European metaweb). The
IUCN names were, as previously, reconciled against GBIF to have an exact
match to the taxonomy.

@@

The method we develop is, ecologically speaking, a ``black box,''
\emph{i.e.} an algorithm that can be understood mathematically, but
whose component parts are not always directly tied to ecological
processes. There is a growing realization in machine learning that
(unintentional) black box algorithms are not necessarily a bad thing
Holm (2019), as long as their constituent parts can be examined (which
is the case with our method). But more importantly, data hold more
information that we may thought; as such, even algorithms that are
disconnected from the model can make correct guesses most of the time
(Halevy \emph{et al.} 2009); in fact, in an instance of ecological
forecasting of spatio-temporal systems, model-free approaches
(\emph{i.e.} drawing all of their information from the data)
outperformed model-informed ones (Perretti \emph{et al.} 2013).

@@

The entire pipeline is implemented in \emph{Julia} 1.6 (Bezanson
\emph{et al.} 2017) and is available under the permissive MIT License at
\href{https://osf.io/2zwqm/}{\texttt{https://osf.io/2zwqm/}}. The
taxonomic cleanup steps are done using \texttt{GBIF.jl}
(\textbf{Dansereau2021SimJl?}). The network embedding and analysis is
done using \texttt{EcologicalNetworks.jl} (Poisot \emph{et al.} 2019;
Banville \emph{et al.} 2021). The phylogenetic simulations are done
using \texttt{PhyloNetworks.jl} (Solís-Lemus \emph{et al.} 2017) and
\texttt{Phylo.jl} (Reeve \emph{et al.} 2016). A complete
\texttt{Project.toml} file specifying the full tree of dependencies is
available alongside the code. This material also includes a fully
annotated copy of the entire code required to run this project
(describing both the intent of the code and discussing some technical
implementation details), a vignette for every step of the process, and a
series of Jupyter notebooks with the text and code. The pipeline can be
executed on a laptop in a matter of minutes, and therefore does not
require extensive computational power.

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy \emph{et al.} 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer 2016). In addition, recent advances show that the
latent variables produced this way can be used to predict \emph{de novo}
network edges. Interstingly, the latent variables do not need to be
prouced by decomposing the network itself; in a recent contribution,
Runghen \emph{et al.} (2021) show that deep artificial neural networks
are able to reconstruct the left and right subspaces of a RDPG, in order
to predict human movement networks from individual/location metadata.
This is an exciting opportunity, as it opens up the possibility of using
additional predictors.

@@

Because RDPG relies on matrix multiplication, the higher dimensions
essentially serve to make specific interactions converge towards 0 or 1;
therefore, for reasonably low ranks, there is no guarantee that the
values in the reconstructed network will be within the unit range. In
order to determine what constitutes an appropriate threshold for
probability, we performed the RDPG approach on the European metaweb, and
evaluated the probability threshold by treating this as a binary
classification problem, specifically assuming that both 0 and 1 in the
European metaweb are all true. Given the methodological details given in
(\textbf{Maiorano2020TetEu?}) and O'Connor \emph{et al.} (2020), this
seems like a reasonable assumption, although one that does not hold for
all metawebs. We used the thresholding approach presented in Poisot
\emph{et al.} (2021b), and picked a cutoff that maximized Youden's \(J\)
statistic (a measure of the informedness (trust) of predictions; Youden
(1950)); the resulting cutoff was 0.22, and gave an accuracy above 0.99.

@@

In order to transfer the knowledge from the European metaweb to the
Canadian species pool, we performed ancestral character estimation using
a Brownian motion model, which is a conservative approach in the absence
of strong hypotheses about the nature of phylogenetic signal in the
network decomposition (Litsios \& Salamin 2012). This uses the estimated
feature vectors for the European mammals to create a state
reconstruction for all species (conceptually something akin to a
trait-based mammalian phylogeny using generality and vulnerability
traits) and allows us to impute the missing (latent) trait data for the
Canadian species that are not already in the European network; as we are
focused on predicting contemporary interactions, we only retained the
values for the tips of the tree. We assumed that all traits (\emph{i.e.}
the feature vectors for the left and right subspaces) were independent,
which is a reasonable assumption as every trait/dimension added to the
t-SVD has an \emph{additive} effect to the one before it. Note that the
Upham \emph{et al.} (2019) tree itself has some uncertainty associated
to inner nodes of the phylogeny. In this case study, we have decided to
not propagate this uncertainty, as it would complexify the process. The
Brownian motion algorithm returns the \emph{average} value of the trait,
and its upper and lower bounds. Because we do not estimate other
parameters of the traits' distributions, we considered that every
species trait is represented as a uniform distribution between these
bounds. The choice of the uniform distribution was made because the
algorithm returns a minimum and maximum point estimate for the value,
and given this information, the uniform distribution is the one with
maximum entropy. Had all mean parameters estimates been positive, the
exponential distribution would have been an alternative, but this is not
the case for the subspaces of an RDPG. In order to examine the
consequences of the choice of distribution, we estimated the variance
per latent variable per node to use a normal distribution; as we show in
Supp. Mat. 2, this decision results in dramatically over-estimating the
number and probability of interactions, and therefore we keep the
discussions in the main text to the uniform case. The inferred left and
right sub-spaces for the Canadian species pool (\(\hat{\mathscr{L}}\)
and \(\hat{\mathscr{R}}\)) have entries that are distributions,
representing the range of values for a given species at a given
dimension.

@@

Dallas \emph{et al.} (2017) suggest that most links in ecological
networks may be cryptic, \emph{i.e.} uncommon or otherwise hard to
observe. This argument essentially echoes Jordano (2016b): the sampling
of ecological interactions is difficult because it requires first the
joint observation of two species, and then the observation of their
interaction. In addition, it is generally expected that weak or rare
links would be more common in networks (Csermely 2004), compared to
strong, persistent links; this is notably the case in food chains,
wherein many weaker links are key to the stability of a system (Neutel
\emph{et al.} 2002). In the light of these observations, the results in
fig.~\ref{fig:inflation} are not particularly surprising: we expect to
see a surge in these low-probability interactions under a model that has
a good predictive accuracy. Because the predictions we generate are by
design probabilistic, then one can weigh these rare links appropriately.
In a sense, that most ecological interactions are elusive can call for a
slightly different approach to sampling: once the common interactions
are documented, the effort required in documenting each rare interaction
may increase exponentially. Recent proposals suggest that machine
learning algorithms, in these situations, can act as data generators
(Hoffmann \emph{et al.} 2019): in this perspective, high quality
observational data can be supplemented with synthetic data coming from
predictive models, which increases the volume of information available
for inference.

@@

As Herbert (1965) rightfully pointed out, ``{[}y{]}ou can't draw neat
lines around planet-wide problems''; in this regard, our approach must
contend with two interesting problems. The first is the limit of the
metaweb to embed and transfer. If the initial metaweb is too narrow in
scope, notably from a taxonomic point of view, the chances of finding
another area with enough related species to make a reliable inference
decrease. This is notably true if the metaweb is assembled in an area
with mostly endemic species. Conversely, the metaweb should be reliably
filled, which assumes that the \(S^2\) interactions in a pool of \(S\)
species have been examined, either through literature surveys or expert
elicitation. The second problem is to determine which area should be
used to infer the new metaweb in, as this determines the species pool
that must be used. In our application, we focused on the mammals of
Canada. The upside of this approach is that information at the country
level is likely to be required by policy makers and stakeholders for
their biodiversity assessment, as each country tends to set goals at the
national level (Buxton \emph{et al.} 2021) for which quantitative
instruments are designed (Turak \emph{et al.} 2017), with specific
strategies often enacted at smaller scales (Ray \emph{et al.} 2021). Yet
these national divisions, in large parts of the world, reflect nothing
except for the legacy of settler colonialism; indeed, the use of
ecological data is not an apolitical act (Nost \& Goldstein 2021).
Operating under the framework of national boundaries, and relying on
data infrastructures designed to answer questions within them, must be
done under the clear realization that this logic contributed to the
ongoing biodiversity crisis (Adam 2014), can reinforce environmental
injustice (Choudry 2013; Domínguez \& Luoma 2020), and on Turtle Island
especially, will probably end up being replaced by Indigenous principles
of land management (Eichhorn \emph{et al.} 2019; No'kmaq \emph{et al.}
2021).

@@

\leavevmode\hypertarget{ref-Arsov2019NetEmb}{}%
Arsov, N. \& Mirceva, G. (2019). \emph{Network Embedding: An Overview}.
Available at: \url{http://arxiv.org/abs/1911.11726}. Last accessed.

@@

\leavevmode\hypertarget{ref-Brose2006ConRes}{}%
Brose, U., Jonsson, T., Berlow, E.L., Warren, P., Banasek-Richter, C.,
Bersier, L.-F., \emph{et al.} (2006). Consumer--Resource Body-Size
Relationships in Natural Food Webs. \emph{Ecology}, 87, 2411--2417.

@@

\leavevmode\hypertarget{ref-Cai2017ComSur}{}%
Cai, H., Zheng, V.W. \& Chang, K.C.-C. (2017). \emph{A Comprehensive
Survey of Graph Embedding: Problems, Techniques and Applications}.
Available at: \url{http://arxiv.org/abs/1709.07604}. Last accessed.

@@

\leavevmode\hypertarget{ref-Cirtwill2019QuaFra}{}%
Cirtwill, A.R., Eklӧf, A., Roslin, T., Wootton, K. \& Gravel, D. (2019).
A quantitative framework for investigating the reliability of empirical
network construction. \emph{Methods in Ecology and Evolution}, 0.

@@

\leavevmode\hypertarget{ref-Csermely2004StrLin}{}%
Csermely, P. (2004). Strong links are important, but weak links
stabilize them. \emph{Trends in Biochemical Sciences}, 29, 331--334.

@@

\leavevmode\hypertarget{ref-Dallas2017PreCry}{}%
Dallas, T., Park, A.W. \& Drake, J.M. (2017). Predicting cryptic links
in host-parasite networks. \emph{PLOS Computational Biology}, 13,
e1005557.

@@

\leavevmode\hypertarget{ref-Grunig2020CroFor}{}%
Grünig, M., Mazzi, D., Calanca, P., Karger, D.N. \& Pellissier, L.
(2020). Crop and forest pest metawebs shift towards increased linkage
and suitability overlap under climate change. \emph{Communications
Biology}, 3, 1--10.

\leavevmode\hypertarget{ref-Halevy2009UnrEff}{}%
Halevy, A., Norvig, P. \& Pereira, F. (2009). The Unreasonable
Effectiveness of Data. \emph{IEEE Intelligent Systems}, 24, 8--12.

@@

\leavevmode\hypertarget{ref-Herbert1965Dun}{}%
Herbert, F. (1965). \emph{Dune}. 1st edn. Chilton Book Company,
Philadelphia.

\leavevmode\hypertarget{ref-Hoffmann2019MacLea}{}%
Hoffmann, J., Bar-Sinai, Y., Lee, L.M., Andrejevic, J., Mishra, S.,
Rubinstein, S.M., \emph{et al.} (2019). Machine learning in a
data-limited regime: Augmenting experiments with synthetic data uncovers
order in crumpled sheets. \emph{Science Advances}, 5, eaau6792.

\leavevmode\hypertarget{ref-Holm2019DefBla}{}%
Holm, E.A. (2019). In defense of the black box. \emph{Science}, 364,
26--27.

@@

\leavevmode\hypertarget{ref-Maiorano2020DatTet}{}%
Maiorano, L., Montemaggiori, A., Ficetola, G.F., O'Connor, L. \&
Thuiller, W. (2020). Data from: Tetra-EU 1.0: A species-level trophic
meta-web of European tetrapods.

@@

\leavevmode\hypertarget{ref-Neutel2002StaRea}{}%
Neutel, A.-M., Heesterbeek, J.A.P. \& de Ruiter, P.C. (2002). Stability
in Real Food Webs: Weak Links in Long Loops. \emph{Science}, 296,
1120--1123.

@@

\leavevmode\hypertarget{ref-Nost2021PolEco}{}%
Nost, E. \& Goldstein, J.E. (2021). A political ecology of data.
\emph{Environment and Planning E: Nature and Space}, 25148486211043503.

\leavevmode\hypertarget{ref-OConnor2020UnvFoo}{}%
O'Connor, L.M.J., Pollock, L.J., Braga, J., Ficetola, G.F., Maiorano,
L., Martinez‐Almoyna, C., \emph{et al.} (2020). Unveiling the food webs
of tetrapods across Europe through the prism of the Eltonian niche.
\emph{Journal of Biogeography}, 47, 181--192.

@@

\leavevmode\hypertarget{ref-Perretti2013ModFor}{}%
Perretti, C.T., Munch, S.B. \& Sugihara, G. (2013). Model-free
forecasting outperforms the correct mechanistic model for simulated and
experimental data. \emph{Proceedings of the National Academy of
Sciences}, 110, 5253--5257.

@@

@@

\leavevmode\hypertarget{ref-Poisot2015SpeWhy}{}%
Poisot, T., Stouffer, D.B. \& Gravel, D. (2015). Beyond species: Why
ecological interaction networks vary through space and time.
\emph{Oikos}, 124, 243--251.
